### 2019-10-31 tex/Beff.tex
big commit

Baldwin (1977) was the first to discover an anti- correlation between the equivalent width (EW) of the \civ $\lambda$1549 emission line and its nearby continuum luminosity in the QSO rest frame (i.e. the Baldwin effect; see a review by Shields 2007).

Explains include:
\begin{itemize}

\item One promising interpretation is the softening of the spectral energy distribution (SED) for increasing luminosity, which lowers the ion (e.g. Netzer et al. 1992; Dietrich et al. 2002).

\item For the Palomar–Green (PG) sample of QSOs, Baskin \& Laor (2004) have found a strong correlation between the \civ EW and $L_{\rm Bol} /L_{\rm} Edd$ , and they have suggested that LBol /LEdd is the primary physical parameter driving the Baldwin effect.

\end{itemize}