### 2019-10-31 tex/CIV_AtomicLineList.tex

% Atomic Line List version: 2.05b21   Constructed: 2017-11-07 13:41 GMT

% Request: 1

% Wavelength range: 1540 - 1560   Unit: Angstrom   Type: Vacuum
% Radial velocity: 0 km/s
% Wavelength accuracy upper limit: 5%
% Element/Spectrum: C  IV
% Minimum line strength: no restrictions
% Include lines without atomic data: true
% Minimum abundance: no minimum
% Lower level energy range: no restrictions   Unit: cm^-1
% Upper level energy range: no restrictions
% Maximum for principal quantum number n: no restrictions
% Transition types included: all
% Transitions from auto-ionizing levels: included

\begin{table*}
\begin{tabular}{rrr@{ -- }lr@{ -- }lr@{ -- }llrrr@{ -- }rl}
\hline
\multicolumn{1}{c}{$\lambda_{\rm lab}^{\rm obs}$\phantom{0000}} & \multicolumn{1}{c}{$\Delta\lambda_{\rm lab}^{\rm obs}$} & \multicolumn{2}{c}{Configurations} & \multicolumn{2}{c}{Terms} & $J_i$ & $J_k$ & \multicolumn{1}{c}{$A_{\rm ki}$} & \multicolumn{1}{c}{$\log(gf)$} & TP flags & \multicolumn{2}{c}{Levels} & Refs \\
\multicolumn{1}{c}{\AA\phantom{0000}} & \multicolumn{1}{c}{\AA} & \multicolumn{2}{c}{} & \multicolumn{2}{c}{} & \multicolumn{2}{c}{} & \multicolumn{1}{c}{s$^{-1}$} & & & \multicolumn{2}{c}{cm$^{-1}$} & \\
\hline
   1548.203\phantom{0000000} & $2.7-2$ &  $1s^{2}2s$ & $1s^{2}2p$  &       $^{2}{\rm S}$ & $^{2}{\rm P}^\circ$ &  $\frac{1}{2}$ & $\frac{3}{2}$  & $2.65+8$ & $ -0.4198$ &     8 &      $0.00$ &  $64591.00$ & 045 \\
   1550.777\phantom{0000000} & $2.7-2$ &  $1s^{2}2s$ & $1s^{2}2p$  &       $^{2}{\rm S}$ & $^{2}{\rm P}^\circ$ &  $\frac{1}{2}$ & $\frac{1}{2}$  & $2.63+8$ & $ -0.7217$ &     8 &      $0.00$ &  $64483.80$ & 045 \\
\hline
\end{tabular}
\end{table*}

\subsection{Carbon IV}
Table~\ref{tab:civ_configs} from 
Atomic Line List version: 2.05b21  
%%Version 2.05b21 of the line list was created: Nov 7, 2017.
\citep{Tunklev1997} \\

\citep[NIST; ][]{Reader2012AAS, Kramida2018}\\

%Wavelength range: 0 - inf   Unit: Angstrom   Type: Vacuum \\
%Radial velocity: 0 km/s\\
%Element/Spectrum: C  IV  \\
%
%-LAB-WAVL-ANG-VAC-|-SPC-|TT|-TERM--|-J\_i-J\_k-|-LVL-EN--CM-1--|-REF---| \\
 % 1548.203         C IV E1  2S-2Po 1/2 - 3/2 0.00 - 64591.00 045 \\
 % 1550.777         C IV E1  2S-2Po 1/2 - 1/2 0.00 - 64483.80 045 \\
%
%  gk*Aki weighted average wavelength:    1549.06    \\
%

\href{https://physics.nist.gov/cgi-bin/ASD/lines1.pl?spectra=C+IV&limits_type=0&low_w=&upp_w=&unit=1&submit=Retrieve+Data&de=0&format=0&line_out=0&en_unit=0&output=0&bibrefs=1&page_size=15&show_obs_wl=1&show_calc_wl=1&unc_out=1&order_out=0&max_low_enrg=&show_av=2&max_upp_enrg=&tsb_value=0&min_str=&A_out=0&intens_out=on&max_str=&allowed_out=1&forbid_out=1&min_accur=&min_intens=&conf_out=on&term_out=on&enrg_out=on&J_out=on}{NIST Atomic Spectra Database Lines Data; C IV: 228 Lines of Data Found}. \\

\href{https://physics.nist.gov/cgi-bin/ASD/energy1.pl?encodedlist=XXT2&de=0&spectrum=C+IV&submit=Retrieve+Data&units=0&format=0&output=0&page_size=15&multiplet_ordered=0&conf_out=on&term_out=on&level_out=on&unc_out=1&j_out=on&lande_out=on&perc_out=on&biblio=on&temp=}{C IV   91 Levels Found
Z = 6, Li isoelectronic sequence}. \\

%%
Kramida, A., Ralchenko, Yu., Reader, J., and NIST ASD Team (2018). NIST Atomic Spectra Database (ver. 5.6.1), [Online]. Available: https://physics.nist.gov/asd [2019, March 22]. National Institute of Standards and Technology, Gaithersburg, MD. DOI: https://doi.org/10.18434/T4W30F

%% 
%% http://www-personal.umich.edu/~cowley/ionen.htm
%% https://en.wikipedia.org/wiki/Ionization_energies_of_the_elements_(data_page)

%%\newpage
%%\input{CIV_AtomicLineList_full}
%%\newpage

%%\newpage
%%\input{CIV_AtomicLineList_mid}
%%\newpage

%%\newpage
%%\input{CIV_AtomicLineList}
%\newpage

\begin{table*}
\begin{tabular}{rrr@{ -- }lr@{ -- }lr@{ -- }llrrr@{ -- }rl}
\hline
\multicolumn{1}{c}{$\lambda_{\rm lab}^{\rm obs}$} & \multicolumn{1}{c}{$\Delta\lambda_{\rm lab}^{\rm obs}$} & \multicolumn{2}{c}{Configurations} & \multicolumn{2}{c}{Terms} & $J_i$ & $J_k$ & \multicolumn{1}{c}{$A_{\rm ki}$} & \multicolumn{1}{c}{$\log(gf)$} & TP flags & \multicolumn{2}{c}{Levels}  \\
\multicolumn{1}{c}{\AA } & \multicolumn{1}{c}{\AA} & \multicolumn{2}{c}{} & \multicolumn{2}{c}{} & \multicolumn{2}{c}{} & \multicolumn{1}{c}{s$^{-1}$} & & & \multicolumn{2}{c}{cm$^{-1}$} & \\
\hline
   1548.203  & $2.7-2$ &  $1s^{2}2s$ & $1s^{2}2p$  &       $^{2}{\rm S}$ & $^{2}{\rm P}^\circ$ &  $\frac{1}{2}$ & $\frac{3}{2}$  & $2.65+8$ & $ -0.4198$ &     8 &      $0.00$ &  $64591.00$  \\
   1550.777  & $2.7-2$ &  $1s^{2}2s$ & $1s^{2}2p$  &       $^{2}{\rm S}$ & $^{2}{\rm P}^\circ$ &  $\frac{1}{2}$ & $\frac{1}{2}$  & $2.63+8$ & $ -0.7217$ &     8 &      $0.00$ &  $64483.80$  \\
\hline
\end{tabular}
\label{tab:civ_configs}
\end{table*}