### 2019-11-07 tex/other_tex/201909/CIV_notes_20190917v1.tex

\documentclass[a4paper,fleqn,usenatbib]{mnras}

\input{format}

\title[High-redshift CLQs]{The first high-redshift Changing Look Quasars}

\author[Bercow]
{The RH John~S.~Bercow, MP, {\it et al.} 
%Nicholas~P.~Ross$^{1}$\thanks{E-mail: npross@roe.ac.uk},    
%K. E. Saavik Ford$^{2,3,4}$,  Matthew Graham$^{5}$,  Barry McKernan$^{2,3,4}$,  
%\newauthor Daniel Stern$^{6}$, 
%Aaron M. Meisner$^{7,8}$, Roberto J. Assef$^{9}$, Arjun Dey$^{10}$, Andrew J. Drake$^{11}$, \newauthor Hyunsung D. Jun$^{12}$, Dustin Lang$^{13,14,15}
\\
% List of institutions
$^{1}$Speaker's Chair, The House of Commons, London, SW1A 0AA \\
%$^{1}$Institute for Astronomy, University of Edinburgh, Royal Observatory, Blackford Hill, Edinburgh EH9 3HJ, United Kingdom \\
%$^{2}$Department of Science, BMCC, City University of New York, New York, NY 10007, USA \\
%$^{3}$Department of Astrophysics, Rose Center for Earth and Space, American Museum of Natural History, Central Park West at 79th Street, NY 10024, USA \\
%$^{4}$Graduate Center, City University of New York, 365 5th Avenue, New York, NY 10016, USA\\
%$^{5}$Cahill Center for Astronomy and Astrophysics, California Institute of Technology, Mail Code 249/17, 1200 E California Blvd, Pasadena CA 91125, USA\\
%$^{6}$Jet Propulsion Laboratory, California Institute of Technology, 4800 Oak Grove Drive, Mail Stop 169-221, Pasadena, CA 91109, USA \\
}

\date{Accepted XXX. Received YYY; in original form ZZZ}
\pubyear{2018}

%\hypersetup{draft}
\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

\begin{abstract}
This is two hundred words, the suggested maximum for a MNRAS
Letter. Lorem ipsum dolor sit amet, autem semper eam ei. Eros
maluisset posidonium eu mea, atqui aperiam nostrum cum ad. Te vocent
volutpat petentium vix, pro senserit convenire vituperatoribus ex, ad
his mucius tamquam appetere. An mea omnium voluptaria scribentur, te
verear iisque vulputate sit. Vim malis omittam probatus cu, vide
tempor ius ex, vim eu case magna rebum. Vis id clita eruditi, in sea
evertitur dissentiunt, per eu deseruisse mnesarchum deterruisset. Eu
tale tempor sea, te mundi congue detracto est. Has ipsum animal timeam
no. Atqui novum nobis et his, pericula interesset mel ut. Ea pro eros
everti legendos. Vel iusto legere euismod id, ea erat debitis
democritum usu. Ut paulo aliquid vix, disputationi delicatissimi
signiferumque ut eam. Utroque impedit ad his. Ei mea sale tritani, ex
eum duis epicuri, at per nonumy delectus electram. Dolorem patrioque
prodesset vis at, sint tritani mel te. Ut nulla solet nonumes has, mel
diam scripta delectus in, ne duo laudem appetere assueverit. An vim
reque munere, labores apeirian his cu. Ei clita placerat praesent nam,
in affert sapientem est. Qui te regione deterruisset. Nam at quodsi
placerat. Pri no quando placerat.
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
accretion, accretion discs -- surveys -- quasars: general -- quasars: individual: J1100-0053 
\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%
%%   SECTION 1  SECTION 1  SECTION 1  SECTION 1  SECTION 1  SECTION 1  
%%   SECTION 1  SECTION 1  SECTION 1  SECTION 1  SECTION 1  SECTION 1  
%%   SECTION 1  SECTION 1  SECTION 1  SECTION 1  SECTION 1  SECTION 1  
%%
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Luminous AGN, i.e. quasars, are now seen to significantly vary their energy output on timescales of weeks to months. 
This observational fact, and the subsequent mismatch in the expected timescale, the ``viscous'' timescale, 
which for a $\approx$10$^{7}$ M$_{\odot}$ is $\sim$hundreds of years, was noted over 30 years ago \citep[e.g.][]{Alloin1985}. 
However, with new photometric light-curve and repeat spectroscopic data, the desire for a deeper understanding 
of AGN accretion disk physics  has been re-established \citep[e.g.][]{Lawrence2018, Antonucci2018}.

Samples of over 100 ``Changing Look'' quasars (CLQs) or ``Changing State'' quasars (CSQs) have now been 
assembled \citep[e.g.][]{MacLeod2019, Graham2019}. The community uses both these
terms as a cover for the underlying physics. For sake of argument, ``Changing Look'' quasars can be 
thought of as the extension of ``regular'' quasar variability \citep[e.g.][]{MacLeod2012} whereas 
the ``Changing State'' quasars (CSQs) have a `state-transition' similar to that in Galactic X-ray binaries
\citep[e.g][]{NodaDone2018, Ruan2019}. 
%
These CLQs have primarily been defined according to their Balmer emission line properties with 
particular attention paid to the H$\beta$ emission line, usually observed from optical spectroscopy. 
As such, current CLQ studies have been at redshifts $z<1$.

While there have been a slew of studies on triply ionized carbon, i.e \civ, broad absorption line quasars
\citep[BAL QSOs; see e.g. Table 1][]{Hemler2019}, dramatic changes in the 
broad {\it emission} line (BEL) of \civ, and indeed, \ciii have not to this point been seen. 
%%

I also don’t think this is a recombination (for H, He) vs. collisionally excited lines (MgII, C IV etc.) thing, though I’m not sure.

Here, we report on three objects 
SDSS J120544.7+342252.4  (hereafter J1205+3422), 
SDSS J163852.93+282707.7 (hereafter J1638+2827) and
SDSS J222818.76+220102.9 (hereafter J2228+2201)  
which show dramatic changes in the \civ broad emission line
properties, as well as in the underlying continuum. We claim these are
the first examples of ``Changing Look Quasars'' at high ($z>1$)
redshift and for emission lines with high ionization potentials
(I.P.'s $>$13.6 eV).

For J1638+2827, over the course of 1279 days observed, 400 days in the rest-frame, the  broad \civ and \ciii BEL start to fade.  
For J2228+2201, over the course of 771 days observed, 240 days in the rest-frame, broad \civ and \ciii BELs both emerge and the 
standard UV/blue continuum slope increases in flux. 
the UV/blue continuum diminishes and the shape of \lya changes.

In this paper we use the wavelengths of 1548.2 and 1550.8 \AA for the \civ doublet
%Example of how to reference these results:
using the values from the National Institute of Standards and Technology (NIST) Atomic Spectra Database 
\citep{Kramida2018}. 
% A., Ralchenko, Yu., Reader, J., and NIST ASD Team (2018). NIST Atomic Spectra Database (ver. 5.6.1), [Online]. Available: https://physics.nist.gov/asd [2019, September 17]. National Institute of Standards and Technology, Gaithersburg, MD. DOI: https://doi.org/10.18434/T4W30F

%The singly ionized hydrogen atom H+ is referred to as HII, doubly ionized oxygen O++ is called OIII, etc.
%11.3 eV is the energy needed for C I to get rid of one electron, and become singly-ionised C II”
%“24.4 eV is energy needed for C I to get rid of two electrons, and become doubly-ionised C III”
%% https://en.wikipedia.org/wiki/Ionization_energies_of_the_elements_(data_page)
Using the \href{https://physics.nist.gov/PhysRefData/ASD/ionEnergy.html}{NIST Atomic Spectra Database Ionization Energies Form}
for ionisation energies, we note 47.89 eV  (3.519 Ry) is the energy required to for neutral C to dislodge three electrons and become triply--ionised \civ, 
and where 64.49 eV (4.74 Ry) is the energy needed to ionize \civ itself.

\subsection{Literature Search}
\citet{Gregory1982} reported time variability in both Lya and \civ 1550 emission lines in the Seyfert galaxy NGC 5548.

From \citet{VandenBerk2001} 
The rest positions of emission lines in quasar spectra, especially the
high-ionization broad lines, are known to vary from their nominal
laboratory wavelengths \citep{Gaskell1982, Wilkes1986,  Espey1989, Espey1990Erratum, 
ZhengSulentic1990, Corbin1990, Weymann1991, TytlerFan1992, 
Brotherton1994, Laor1995, McIntosh1999}.

\citet{Wilhite2006} examine the variability of the high-ionizaton
C~\textsc{iv}$\lambda{1549}$ line in a sample of 105 quasars observed
at multiple epochs by the Sloan Digital Sky Survey.  We find a strong
correlation between the change in the C~\textsc{iv} line flux and the
change in the line width, but no correlations between the change in
flux and changes in line center and skewness.  The relation between
line flux change and line width change is consistent with a model in
which a broad line base varies with greater amplitude than the line
core. The objects studied here are more luminous and at higher
redshift than those normally studied for variability, ranging in
redshift from 1.65 to 4.00 and in absolute $r$-band magnitude from
roughly $-$24 to $-$28.  Using moment analysis line-fitting
techniques, we measure line fluxes, centers, widths and skewnesses for
the C~\textsc{iv} line at two epochs for each object.  The well-known
Baldwin Effect is seen for these objects, with a slope $\beta =
-0.22$.  The sample has a median intrinsic Baldwin Effect slope of
$\beta_{int} = -0.85$; the C~\textsc{iv} lines in these
high-luminosity quasars appear to be less responsive to continuum
variations than those in lower luminosity AGN.  Additionally, we find
no evidence for variability of the well known blueshift of the
C~\textsc{iv} line with respect to the low-ionization
Mg~\textsc{ii}$\lambda$2798 line in the highest flux objects,
indicating that this blueshift might be useful as a measure of
orientation.

%% NAZAROVA 2003
%% BROAD EMISSION LINES La, C IV AND Hb IN NGC 5548
%% Astronomical and Astrophysical Transactions
%% Vol. 22, Nos. 4–5, August–October 2003, pp. 681–689

%% Punsly 2010
%% Astrophysical Journal, 713, 232
%% THE REDSHIFTED EXCESS IN QUASAR C iv BROAD EMISSION LINES

\citet{Denney2015}, 
%Denney, Kelly; SDSS-RM Team, ``The Effects of S/N on Measuring CIV Broad Emission Line Widths in Quasars - An Early Science Result from the Sloan Digital Sky Survey Reverberation Mapping Project'' ; http://adsabs.harvard.edu/abs/2015AAS...22520403D
The Sloan Digital Sky Survey Reverberation Mapping (SDSS-RM) Project
spectroscopically monitored $\sim$850 quasars over a seven square
degree field for approximately six months with the BOSS
spectrograph. While the time series analysis of the quasar continuum
and emission-line variability, and thereby the potential to measure
reverberation time delays from this sample, is still underway, we have
used the subset (roughly half) of this sample containing the \civ 1549\AA 
broad emission line to investigate the effects of S/N on measuring
this broad emission line width. Line width measurements are necessary
for inferring the broad line region gas velocities used to determine
the quasar black hole mass. However, literature studies have suggested
line width measurements, and therefore black hole mass estimates, may
be biased by the use of relatively low S/N, typical survey-quality,
data, compared to the high S/N data that calibrate single-epoch black
hole mass scaling relations. The SDSS-RM data set provides a unique
opportunity to search for any systematic uncertainties in CIV line
width measurements through a comparison of line widths measured from
the single-epoch spectra, which are roughly equivalent to other
SDSS/BOSS quasar spectra (though still twice the exposure time) to
those measured from the high-S/N, campaign co-added spectra, which
contain more than 30 epochs of each source. Here we present results
from this investigation, which have implications for estimating CIV
single-epoch black hole masses from the SDSS/BOSS and other surveys.

%% Coatman et al. (2017, AAS; http://adsabs.harvard.edu/abs/2017AAS...22930203C) 
%%Accurate black-hole (BH) mass estimates for high-redshift (z>2) quasars are essential for better understanding the relationship between super-massive BH accretion and star formation. Progress is currently limited by the large systematic errors in virial BH-masses derived from the CIV broad emission line, which is often significantly blueshifted relative to systemic, most likely due to outflowing gas in the quasar broad-line region. We have assembled Balmer-line based BHmasses for a large sample of 230 high-luminosity (1045.5-1048 ergs-1), redshift 1.5<z<4 quasars, which, for the first time, span the entire range of CIV blueshifts seen in the quasar population. We find the CIV-based BH-masses to be larger than the corresponding Balmer line-based masses by almost an order of magnitude at the most extreme blueshifts (˜5000 kms-1). An empirical correction to the CIV BH-masses is derived, which depends only on the properties of the CIV line itself (i.e. blueshift and FWHM). 
%%
\citet{Coatman2017AAS} show that this new correction now enables the derivation of un-biased CIV-based virial BH masses for the majority of high-luminosity, high-redshift quasars.In the same high-luminosity quasar sample, we find the narrow [OIII] emission to be weaker and more asymmetric than is generally found in  lower-luminosity AGN and that a significant fraction of our quasars have exceptionally broad (FWHM $> 3000 km s^{-1}$), blueshifted [OIII] emission. We find a strong correlation between the \civ and [OIII] blueshifts. This correlation holds even for quasars at fixed
luminosity and suggests that broad line region outflows in quasars are
connected to galaxy-scale winds.

\citet{Coatman2016} show that %%Black hole masses are crucial to understanding the physics of the connection between quasars and their host galaxies and measuring cosmic black hole-growth. 
at high redshift, $z>2.1$, black hole masses are normally derived using the velocity width of the C IV λλ1548, 1550 broad emission line, based on the assumption that the observed velocity widths arise from virial- induced motions. In many quasars, the C IV emission line exhibits significant blue asymmetries (‘blueshifts’) with the line centroid displaced by up to thousands of km s−1 to the blue. These blueshifts almost certainly signal the presence of strong outflows, most likely originating in a disc wind. We have obtained near-infrared spectra, including the Hα λ6565 emission line, for 19 luminous (LBol = 46.5–47.5 erg s−1) Sloan Digital Sky Survey quasars, at redshifts 2 < z < 2.7, with C IV emission lines spanning the full range of blueshifts present in the population. A strong correlation between C IV velocity width and blueshift is found and, at large blueshifts, >2000kms−1, the velocity widths appear to be dominated by non-virial motions. Black hole masses, based on the full width at half-maximum of the C IV emission line, can be overestimated by a factor of 5 at large blueshifts. A larger sample of quasar spectra with both C IV and H β, or Hα, emission lines will allow quantitative corrections to C IV-based black hole masses as a function of blueshift to be derived. We find that quasars with large C IV blueshifts possess high Eddington luminosity ratios and that the fraction of high-blueshift quasars in a flux-limited sample is enhanced by a factor of approximately 4 relative to a sample limited by black hole mass.

%% \citet{Coatman2017} 
The \civ $\lambda \lambda$1498,1501 broad emission line is visible in optical spectra to redshifts exceeding $z\sim5$. \civ has long been known to exhibit significant displacements to the blue and these `blueshifts' almost certainly signal the presence of strong outflows. As a consequence, single-epoch virial black hole (BH) mass estimates derived from \civ velocity widths are known to be systematically biased compared to masses from the hydrogen Balmer lines. 
\citet{Coatman2017} use a large sample of 230 high-luminosity (LBol = 1045.5-1048 erg s-1), redshift $1.5 < z < 4.0$ quasars with both \civ and Balmer line spectra, we have quantified the bias in \civ BH masses as a function of the \civ blueshift. \civ BH masses are shown to be a factor of 5 larger than the corresponding Balmer-line masses at C IV blueshifts of 3000 km s-1 and are overestimated by almost an order of magnitude at the most extreme blueshifts, $\gtrsim$5000 km s$^{-1}$. Using the monotonically increasing relationship between the \civ blueshift and the mass ratio BH(\civ)/BH(H$\alpha$), we derive an empirical correction to all C IV BH masses. The scatter between the corrected C IV masses and the Balmer masses is 0.24 dex at low C IV blueshifts ($\sim$0 km s-1) and just 0.10 dex at high blueshifts ($\sim$3000 km s$^{-1}$), compared to 0.40 dex before the correction. The correction depends only on the C IV line properties - i.e. full width at half-maximum and blueshift - and can therefore be applied to all quasars where C IV emission line properties have been measured, enabling the derivation of unbiased virial BH-mass estimates for the majority of high-luminosity, high-redshift, spectroscopically confirmed quasars in the literature.

\citet{Sun2018} use the multi-epoch spectra of 362 quasars from the Sloan Digital Sky Survey Reverberation Mapping project to investigate the dependence of the blueshift of \civ relative to \mgii on quasar properties. We confirm that high-blueshift sources tend to have low \civ equivalent widths (EWs), and that the low-EW sources span a range of blueshift. Other high-ionization lines, such as He II, also show similar blueshift properties. The ratio of the line width (measured as both the full width at half maximum and the velocity dispersion) of \civ to that of \mgii increases with blueshift. Quasar variability enhances the connection between the \civ blueshift and quasar properties (e.g., EW). The variability of the \mgii line center (i.e., the wavelength that bisects the cumulative line flux) increases with blueshift. In contrast, the C IV line center shows weaker variability at the extreme blueshifts. Quasars with the high-blueshift \civ lines tend to have less variable continuum emission, when controlling for EW, luminosity, and redshift. Our results support the scenario that high-blueshift sources tend to have large Eddington ratios.

%% Romain A. Meyer, Sarah E. I. Bosman, Richard S. Ellis  https://arxiv.org/abs/1902.04558v1
\citet{Meyer2019} present the results of a model-independent investigation of the rest-frame UV spectra from a comprehensive sample of 394 quasars in the redshift range $1.5\leq z \leq 7.5$. We fit the main Broad Emission Lines (BELs) in the rest-frame range 1280 \AA $\leq \lambda \leq $3000 \AA (O I, C II, Si IV, C III, C IV and Mg II) with a lightly-supervised spline fitting technique. Redshifts are derived from the peaks of each fitted BEL and used to compute relative velocity shifts. We show that our method gives unbiased velocity shifts and is insensitive to spectral resolution and instrumental parameters. It is found that the average blueshift of the \civ\, line with respect to several low-ionisation lines in luminosity-matched samples does not significantly evolve over $1.5\leq z\leq 6$. However, the average blueshift increases significantly by a factor $\sim$2.5 at $z \geq 6$. We propose that this redshift evolution can be explained by \civ\, winds launched perpendicularly to an accretion disk with increased torus opacity at high-redshift, coupled with a potential orientation-driven selection bias. Our results open new exciting avenues of investigation into young quasars in the reionisation epoch.

\citet{Richards2011} 
use a sample of $\sim$30,000 quasars from the 7th Data Release of the Sloan Digital Sky Survey, and explore the range of properties exhibited by high-ionization, broad emission lines, such as \civ $\lambda$1549. Specifically, we investigate the anti-correlation between continuum luminosity and emission-line equivalent width (the Baldwin Effect (BEff)) and the ``blueshifting''of the high-ionization emission lines with respect to low-ionization emission lines. 
%Employing improved redshift determinations from Hewett \& Wild, the blueshift of the C IV emission line is found to be nearly ubiquitous, with a mean shift of ~810 km s-1 for radio-quiet (RQ) quasars and ~360 km s-1 for radio-loud (RL) quasars. The BEff is present in both RQ and RL samples. We consider these phenomena within the context of an accretion disk-wind model that is modulated by the nonlinear correlation between ultraviolet and X-ray continuum luminosity. Composite spectra are constructed as a function of C IV emission-line properties in an attempt to reveal empirical relationships between different line species and the continuum. Within a two-component disk+wind model of the broad emission-line region (BELR), where the wind filters the continuum seen by the disk component, we find that RL quasars are consistent with being dominated by the disk component, while broad absorption line quasars are consistent with being dominated by the wind component. Some RQ objects have emission-line features similar to RL quasars; they may simply have insufficient black hole (BH) spin to form radio jets. Our results suggest that there could be significant systematic errors in the determination of L bol and BH mass that make it difficult to place these findings in a more physical context. 
It is possible to classify quasars in a paradigm where the diversity of BELR parameters is due to differences in an accretion disk wind between quasars (and over time); these differences are underlain primarily by the spectral energy distribution, which ultimately must be tied to BH mass and accretion rate.

Also from \citet{Richards2011}, 
At high redshift, two well-known properties of the Civ emission line are evident in quasar spectra (see Section 4). An anti-correlation between luminosity and the equivalent width (EQW) of C iv, generally known as the Baldwin Effect (BEff; Baldwin 1977), was first identified in 20 (mostly radio-loud (RL)) quasars and has been confirmed in much larger samples (e.g., Wu et al. 2009). The other effect is the blueshifting of the C iv with respect to the systemic redshift (Gaskell 1982; Wilkes 1984). As with the BEff, these blueshifts have not only been confirmed with larger samples, but appear to be nearly ubiquitous in the quasar population (Sulentic et al. 2000b; Richards et al. 2002a).

%\includegraphics\includegraphics[]{../../../../../Users/npr1/Downloads/Richards_2011_AJ_141_167.pdf}

%\citet{1994}
%\includegraphics[]{../../../../../Users/npr1/Downloads/1994ApJ___425__622P.pdf}

%\citet{1994}
%\includegraphics[]{../../../../../Users/npr1/Downloads/1994ASPC___69____1P.pdf}

\citet{Denney2016}
%\includegraphics[]{../../../../../Users/npr1/Downloads/Denney_2016_ApJS_224_14.pdf}
investigate the dependence on data quality of quasar properties
measured from the C IV emission line region at high redshifts.  Our
measurements come from 32 epochs of Sloan Digital Sky Survey
Reverberation Mapping Project spectroscopic observations of 482 $z >
1.46$ quasars. We compare the differences between measurements made
from the single-epoch (SE) and coadded spectra, focusing on the \civ
$\lambda$1549 emission line because of its importance for studies of
high-redshift quasar demographics and physical properties, including
black hole masses. In addition to statistical errors increasing (by
factors of $\sim2-4$), we find increasing systematic offsets with
decreasing signal-to-noise ratio (S/N).

\citet{Grier2015}....
%\includegraphics[]{../../../../../Users/npr1/Downloads/Grier_2015_ApJ_806_111.pdf}

\citep{Corbin1991, Dimitrijevic1992, Wills1993, Brotherton1994, Osmer1994} ...

\citep{Nazarova2003} ...

%\includegraphics[]{../../../../../Users/npr1/Downloads/Misawa_2007_ApJ_660_152.pdf}
%\includegraphics[]{../../../../../Users/npr1/Downloads/Punsly_2010_ApJ_713_232.pdf}
%\includegraphics[]{../../../../../Users/npr1/Downloads/Sun_2015_ApJ_811_42.pdf}
%\includegraphics[]{../../../../../Users/npr1/Downloads/Joshi_2019_ApJ_871_43.pdf}

\iffalse
\begin{figure}
  \centering
  \includegraphics[width=8.7cm, trim=0.2cm 0.2cm 0.2cm 0.2cm, clip]
  {figures/J2228+2201.png}
  \vspace{-12pt}
  \caption[]{J222818.7+220102.6. 
% EMERGING., 
The light curve shows it brightening from 20.5 to 19.5 over the 771 days observed-frame between the two SDSS spectra available. 
A quasar at $z = 2.2$ with the two spectra showing CIV and CIII] emerging on a new blue contiuum slope.}
  \label{fig:disk_suppression}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=8.7cm, trim=0.2cm 0.2cm 0.2cm 0.2cm, clip]
  {figures/J1638+2827.png}
  \vspace{-12pt}
  \caption[]{The quasar SDSS J163852.9+282708. 
% Disappearing 
A quasar at $z = 2.2$ with the two spectra showing CIV and CIII] fading.}
  \label{fig:disk_suppression}
\end{figure}
\fi

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%   SECTION 2   SECTION 2   SECTION 2   SECTION 2   SECTION 2   SECTION 2  
%%   SECTION 2   SECTION 2   SECTION 2   SECTION 2   SECTION 2   SECTION 2  
%%   SECTION 2   SECTION 2   SECTION 2   SECTION 2   SECTION 2   SECTION 2  
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data}
%\subsection{Selection/Identification}
%\subsection{SDSS spectra}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aliquam porta
sodales est, vel cursus risus porta non. Vivamus vel pretium
velit. Sed fringilla suscipit felis, nec iaculis lacus convallis
ac. Fusce pellentesque condimentum dolor, quis vehicula tortor
hendrerit sed. Class aptent taciti sociosqu ad litora torquent per
conubia nostra, per inceptos himenaeos. Etiam interdum tristique diam
eu blandit. Donec in lacinia libero.

Sed elit massa, eleifend non sodales a, commodo ut felis. Sed id
pretium felis. Vestibulum et turpis vitae quam aliquam convallis. Sed
id ligula eu nulla ultrices tempus. Phasellus mattis erat quis metus
dignissim malesuada. Nulla tincidunt quam volutpat nibh facilisis
euismod. Cras vel auctor neque. Nam quis diam risus.

\begin{table}
 \centering
 \begin{tabular}{l l l l}
  \hline \hline 
% &&\\
   Quantity \ Object                           & J2228+2201     &  J1638+2827 \\
% &&\\
 \hline 
    &&\\
    R.A. / deg                                        &    337.078194      &  249.720559\\
    Declination / deg                            &    +22.017478      &  +28.452159 \\
    redshift, $z$                                    &   2.222$\pm$0.00038   &  2.185$\pm$0.00043          \\
    &&\\ 
    \multirow{2}{*}{Plate, Fiber, MJD}   & 6118, 720, 56189	     &  2948, 614, 54553	  \\
                                         & 7582, 790, 56960	     & 5201, 178, 55832 \\    
 %   &&\\ 
    %$M_{i}(z=2)$  / mag                          &   ?                       & ?                \\
    %log $(L_{\rm bol} / {\rm erg s}^{-1}) $      &   ?             & 45.56$\pm$0.004      & 45.07$\pm$0.004 \\
    %log $(M_{\rm BH} / M_{\odot})  $           &  ?                  & 8.43$\pm$0.03           & 8.46$\pm$0.02 \\
    %Eddington ratio  (\%)                        &  ?                               &  10.7                           &  3.2     \\ 
    &&\\
    \hline \hline 
  \end{tabular}
  \caption{Physical properties of J1100-0053, J2317+0005 and J1052+1519 using the
    methods from \citet{Shen2011}. *This spectrum was used to estimate
    the quantities reported.  We use the regular definition of $L_{\rm
      Edd} = 4 \pi G M m_{\rm p} c /\sigma_{T} =
    1.26\times10^{38}\left (M/M_{\odot} \right )$ erg s$^{-1}$.} 
 \label{tab:Shen_props}
\end{table}

Double Spectrograph (DBSP) is a low-to-medium resolution ($R\sim$1,000
to 10,000) grating instrument for the Palomar 200-inch telescope
Cassegrain focus that uses a dichroic (at, e.g., 5200\AA\ ) to split
light into simultaneously observed separate red and blue channels.

The Low Resolution Imaging Spectrometer (LRIS) is a visible-wavelength
imaging and spectroscopy instrument commissioned in 1993 and operating
at the Cassegrain focus of 10-meter Keck I telescope.

\begin{table*}
 \centering
 \begin{tabular}{r  r  r r r   r r r r}
  \hline \hline 
% &&\\
   Object                         & \multirow{3}{*}{Redshift} & \multirow{3}{*}{Mag}      & \multirow{3}{*}{MJD} & \multirow{3}{*}{Date}  & \multirow{3}{*}{Instrument}   & Exposure    & \multirow{3}{*}{Spectrum} & \multirow{3}{*}{Notes} \\
   R.A. / deg                   &               &               &            &                                &                                                                       &  Time         & & \\
   Decl. / deg                 &               &               &            &                                &                                                                       &  / seconds         & & \\

% &&\\
 \hline 
                                       &                &               &            &                              &                   &                             &  & \\
   J120544.7+342252.4   & 2.071     &   17.99   & 53498 & 2005-May-08   & SDSS             & ?                          & 53498-2089-427  & \\
          181.436164           &                &               & 58538 & 2019-Feb-24     & DBSP            &  2$\times$900    &                               &  Conditions? \\
         +34.381229           &                &               & 58693 & 2019-Jul-29      & DBSP            &  2$\times$1200   &                                &   \\
                                       &                &               &              &                           &                   &                             &  & \\
   J163852.9+282707.7   & 2.182     &   19.77    & 54553 & 2008-Mar-28    & SDSS             & ?                          & 54553-2948-614  & \\
      249.720558407	       &               &              & 55832  & 2011-Sep-28     & BOSS            &   3603                  & 55832-5201-178  & \\
       +28.452158884        &               &              & 58583 & 2019-Apr-10        & LRIS              &  2$\times$900    &   & \\
                                       &                &               &              &                           &                   &                             &  & \\
    J222818.7+220102.9   & 2.222     & 19.97    & 56189 & 2012-Sep-19     & BOSS             &  2702               & 56189-6118-720   & \\
        337.078194              &              &              & 56960 & 2014-Oct-30    & BOSS             & ?                         & 56960-7582-790     & QSO1-REOBS \\ 
        +22.017478                                &               &             & 58693 & 2019-Jul-29      & DBSP              & 2$\times$1200  &       &    \\
                                       &                &               &            &                             &                   &                              &  & \\
%   J232446.7+140029.5   &  0.916     &  18.83   &  56209 & 2012-Oct-09   & BOSS             & ?                             & 56209-6148-743   & \\
 %                                      &                &               &   58662 & 2019-Jun-28  &   (??)DBSP       & ?                            &  & \\
  %                                     &                &               &              &                           &                   &                              &  & \\
   \hline \hline   
  \end{tabular}
  \caption{Spectrum is SDSS/BOSS MJD-Plate-FiberID combination. 
Exposure times are from the {\tt plate.fits} file. 
DBSP on the Palomar 200-inch telescope is the } 
 \label{tab:obs_notes}
\end{table*}

\subsection{Multi-wavelength properties}
MIR \\
Radio \\

%%\newpage
%% For the MNRAS Letter, not sure this can take up a full page...
\begin{figure}
  \centering
  \includegraphics[width=8.7cm, trim=0.2cm 0.2cm 0.2cm 0.2cm, clip]
  {../plots/spectral/The_3_CIV_CLQs_20190917.pdf}
% {figures/The_3_CIV_CLQs_20190917.pdf}
   \vspace{-12pt}
  \caption[]{The threes high-$z$ CLQ quasars; 
SDSS J1205+3422 (top), 
SDSS J1638+2827 (middle), 
SDSS J2228+2201 (bottom). 
%A quasar at $z = 2.2$ with the two spectra showing CIV and CIII] fading.
}
  \label{fig:civ_clqs}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%   SECTION 3   SECTION 3   SECTION 3   SECTION 3   SECTION 3   SECTION 3  
%%   SECTION 3   SECTION 3   SECTION 3   SECTION 3   SECTION 3   SECTION 3  
%%   SECTION 3   SECTION 3   SECTION 3   SECTION 3   SECTION 3   SECTION 3  
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
Figure~\ref{fig:civ_clqs} presents the threes high-$z$ CLQ quasars. 
Nunc semper quam et leo interdum vulputate eu quis magna. Sed nec arcu
at orci egestas convallis. Aenean quam velit, aliquam vitae viverra
in, elementum vel elit. Nunc suscipit aliquet sapien a suscipit. Cras
nulla ipsum, posuere eu fringilla sit amet, dapibus ultricies
nulla. Nullam eu augue id purus mollis dignissim sed et
libero. Phasellus eget justo sed neque pellentesque egestas nec id
arcu. Donec facilisis pulvinar sapien et fringilla. Suspendisse
vestibulum rhoncus sapien id laoreet. Morbi et orci vitae tortor
imperdiet imperdiet. In hac habitasse platea dictumst. Vivamus vel
neque id mi ultrices tristique. Integer quam libero, ornare vel
gravida in, feugiat a ante. Nam dapibus, tellus vitae pellentesque
cursus, dui nisl egestas augue, non fermentum nisl est nec
nisi. Vestibulum nec mi justo, eget dapibus velit.

\begin{figure}
  \centering
  \includegraphics[width=8.7cm, trim=0.2cm 0.2cm 0.2cm 0.2cm, clip]
  {../plots/Wilhite/Wilhite_2006_Fig2_redux_20190830.png}
% {figures/The_3_CIV_CLQs_20190917.pdf}
   \vspace{-12pt}
  \caption[]{
The Change in \civ line width vs. line flux change. 
Data from \citet{Wilhite2006} with a sample of 105 quasars observed at
multiple epochs by the Sloan Digital Sky Survey. We find a strong
correlation between the change in the C iv line flux and the change in
the line width, but no correlations between the change in flux and
changes in line center and skewness.}
  \label{fig:Wilhite2006_comparison}
\end{figure}
From Figure~\ref{fig:Wilhite2006_comparison} there appears to be a
strong correlation between the change in the line flux and the change
in the line width.  Figure~\ref{fig:Wilhite2006_comparison} shows the
epoch-to-epoch flux ratio versus the ratio of line widths.

In the context of CLQs at lower-$z$...  Sociis natoque penatibus et
magnis dis parturient montes, nascetur ridiculus mus. Duis tempus,
lectus nec ultricies mollis, mi orci feugiat nulla, a bibendum velit
orci nec lacus. Duis a odio in nisi egestas dictum. Nullam vel quam
mauris, eget consectetur orci. Morbi ac mi sit amet neque consectetur
tempus ac eget est. Curabitur malesuada arcu sit amet metus dictum at
dapibus arcu accumsan. Fusce sollicitudin luctus rutrum.  Balmer CLQs
vs. \civ CLQs...

\subsection{J1205+3422 }
SDSS J120544.67+342252.4, observed by SDSS, first reported in \citet{Schneider2007}. 
Redshift of $z=$2.068$\pm$0.00027. 
%% targeting_flags	SERENDIP_BLUE QSO_CAP
plate	2089 
mjd	53498
fiberid	427

\subsection{J1638+2827}
SDSS J163852.93+282707.7       16h38m52.9s +28d27m08s  \\
First reported in \citet{Schneider2010}. 
Selected by {\tt SERENDIP\_BLUE} and {\tt QSO\_MAG\_OUTLIER} in SDSS \citep{Richards2002}, meaning...
Selected as a $z>2$ known quasar and passed BOSS quasar target selections \citep{Ross2012}.

R.A. (J2000) = 249.72058 degs. \\
Decl. (J2000) = 28.45217 degs. \\
\href{skyserver.sdss.org/dr15/en/tools/explore/Summary.aspx?id=1237662301375824232}{SDSS SkyServer} 
All Spectra of this Object:: \\
specObjId	plate	MJD	fiber	ra	dec	redshift \\
5855854441601671168	5201	55832	178	2.186$\pm$0.00073\\
3319321776793610240	2948	54553	614	2.185$\pm$0.00043\\

\subsection{J2228+2201}
SDSS J222818.76+220102.9 (R.A. 337.078194926, Decl. 22.017477924) was first reported in 
\citet{Paris2017}.

%%\href{https://skyserver.sdss.org/dr12/en/tools/chart/navi.aspx?ra=337.077916666667&dec=22.0173888888889&scale=0.2&width=120&height=120&opt=
\href{skyserver.sdss.org/dr15/en/tools/explore/summary.aspx?id=1237678579819479655}{SDSS SkyServer}
All Spectra of this Object:: \\
specObjId	plate	MJD	fiber	ra	dec	redshift \\
8536790363997708288	7582	56960	790	2.222$\pm$0.00038\\
6888453645991387136	6118	56189	720	2.217$\pm$0.00208\\

%Plate: 2948, MJD: 54553, Fiber: 0614
%https://dr12.sdss.org/spectrumDetail?mjd=54553&fiber=614&plateid=2948

\begin{table*}
  \centering
  \begin{tabular}{l  l  l    lll}
    \hline \hline 
    Line 	    &   Rest	      & Line       & Line	       &	Line			                 &   Continuum \\ 
    Name	    &   Wavelength  &  $z$       & $\sigma$ 	&	Flux			                 & \\
                    &    $\AA $        &              & /  km s$-1$	&     10$^{-17}$ erg/cm$2$/s &  10$^{-17}$ erg/cm$^2$/s/\AA  \\
Ly$\alpha$  &	 1215.7	&	2.186      &	1362	&	510.5		                &	7.224 \\
\nv		    &   1240.8	&	2.184     &	 854.6	&	122.4	                    	&	5.546 \\ 
\civ		    &  1549.5	&	2.184    &	1927	&	462.6		                &	2.722 \\ 
\heii	    &	 1640.4	&	2.184    &	1927	&    	  8.067	                   	&	2.447 \\  
\ciii 	    &	 1908.7	&	2.184    &	1927	&	164.6	                   	&	1.966  \\ 
\mgii	    &	 2800.3	&	2.184   & 	1927	&	187.4	                   	&	1.513  \\
   \hline \hline   
  \end{tabular}
  \caption{Line Measurement Information}
 \label{tab:line_values}
\end{table*}

\begin{table*}
  \centering
  \begin{tabular}{l  lll lll  }
    \hline \hline 
                                    &                               &                                           &                                                      &                                  &                         & \\  
                                  \multicolumn{7}{c}{J1638+2827}  \\
                                    &                               &                                           &                                                      &                                  &                         & \\  
   \hline                       
%                                   &                               &                                           &                                                      &                                  &                         & \\  
    Line      	             & Line  $\sigma$     & Line  Flux                           &   Continuum                                 & Line  $\sigma$          & Line  Flux        &   Continuum                                    \\
%    Wavelength	             & /  km s$^{-1}$         &   10$^{-17}$ erg/cm$2$/s  & 10$^{-17}$ erg/cm$^2$/s/\AA      &  /  km s$-1$             & / 10$^{-17}$ erg/cm$2$/s   & /10$^{-17}$ erg/cm$^2$/s/\AA       \\
                                     & \multicolumn{3}{l}{MJD 54553}                                                                                   & \multicolumn{3}{l}{MJD 55832}                                                                                                          \\
    Ly$\alpha$1215.7    &  1362	            &	510.5		                  &	7.224                                           &  1769	                &   633.9	              &  5.376     \\
    \nv1240.8	            &   854.6	            &	122.4	                    	  &	5.546                                           & 1012	                &  107.1		      &  4.645\\ 
    \civ1549.5         	    &	 1927	            &	462.6		                  &	2.722                                           & 2287	                &  125.1		      &  1.256      \\  
    \heii1640.4	            &	 1927    	            &      8.067	                   	  &	2.447                                           & 2287	                &    10.68	      &  1.001\\  
    \ciii1908.7        	    &	 1927	            &	164.6	                   	  &	1.966                                          & 2287	 	                &     29.53	      &  0.6008\\ 
    \mgii2800.3	    &   1927	            &	187.4	                   	  &	1.513                                          & 2287 	 	         &    34.66	      &  0.3461  \\
                                    &                               &                                           &                                                      &                                  &                           &                \\  
   \hline \hline   
  \end{tabular}
  \caption{Line Measurement Information. 
    Line $\sigma$ in units of   km s$-1$; 
    Line flux          in units of  10$^{-17}$ erg/cm$^2$/s; 
    Continuum      in units of  10$^{-17}$ erg/cm$^2$/s/ \AA; 
}
 \label{tab:line_values}
\end{table*}

%Plate: 5201, MJD: 55832, Fiber: 178
%https://dr12.sdss.org/spectrumDetail?mjd=55832&fiber=178&plateid=5201

%Line		Rest		        Line	        Line		        Line			   Continuum
%Name		Wavelength	z	        sigma 		Flux	
%                       [Å]			               [km s-1]  	[10-17erg/cm2/s] 	   [10-17erg/cm2/s/Å]
%Ly_alpha	1215.7		2.186	1769		633.9		   5.376
%N_V		1240.8		2.185	1012		107.1		   4.645
%C_IV		1549.5		2.185	2287		125.1		   1.256
%He_II		1640.4		2.185	2287		 10.68		   1.001
%C_III]		1908.7		2.185	2287	 	 29.53		   0.6008
%Mg_II		2800.3		2.185	2287 	 	 34.66		   0.3461

%So, ratios, MJD 54553 / MJD 55832::
%$Ly_alpha	1215.7		—	0.770		0.805			1.344			
%N_V		1240.8		—	0.844		1.143			1.194
%C_IV		1549.5		—	0.843		3.698			2.167			
%He_II		1640.4		—	0.843		0.755			2.445			
%C_III]		1908.7		—	0.843		5.574			3.27			
%Mg_II		2800.3		—	0.843		5.407			4.37

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%   SECTION 4   SECTION 4   SECTION 4   SECTION 4   SECTION 4   SECTION 4  
%%   SECTION 4   SECTION 4   SECTION 4   SECTION 4   SECTION 4   SECTION 4  
%%   SECTION 4   SECTION 4   SECTION 4   SECTION 4   SECTION 4   SECTION 4  
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion and Conclusions}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Aliquam porta
sodales est, vel cursus risus porta non. Vivamus vel pretium
velit. Sed fringilla suscipit felis, nec iaculis lacus convallis
ac. Fusce pellentesque condimentum dolor, quis vehicula tortor
hendrerit sed. Class aptent taciti sociosqu ad litora torquent per
conubia nostra, per inceptos himenaeos. Etiam interdum tristique diam
eu blandit. Donec in lacinia libero.

Sed elit massa, eleifend non sodales a, commodo ut felis. Sed id
pretium felis. Vestibulum et turpis vitae quam aliquam convallis. Sed
id ligula eu nulla ultrices tempus. Phasellus mattis erat quis metus
dignissim malesuada. Nulla tincidunt quam volutpat nibh facilisis
euismod. Cras vel auctor neque. Nam quis diam risus.

Nunc semper quam et leo interdum vulputate eu quis magna. Sed nec arcu
at orci egestas convallis. Aenean quam velit, aliquam vitae viverra
in, elementum vel elit. Nunc suscipit aliquet sapien a suscipit. Cras
nulla ipsum, posuere eu fringilla sit amet, dapibus ultricies
nulla. Nullam eu augue id purus mollis dignissim sed et
libero. Phasellus eget justo sed neque pellentesque egestas nec id
arcu. Donec facilisis pulvinar sapien et fringilla. Suspendisse
vestibulum rhoncus sapien id laoreet. Morbi et orci vitae tortor
imperdiet imperdiet. In hac habitasse platea dictumst. Vivamus vel
neque id mi ultrices tristique. Integer quam libero, ornare vel
gravida in, feugiat a ante. Nam dapibus, tellus vitae pellentesque
cursus, dui nisl egestas augue, non fermentum nisl est nec
nisi. Vestibulum nec mi justo, eget dapibus velit.

Cras in laoreet mauris. Vivamus nec nulla a dui commodo
adipiscing. Proin vulputate lectus nec arcu iaculis sit amet auctor
ligula ultricies. Phasellus condimentum gravida tincidunt. Phasellus
et mauris ac nibh vestibulum vehicula. Morbi et augue id purus gravida
sagittis quis in sem. Phasellus quis risus bibendum eros luctus
auctor.

Etiam mollis viverra nisi eget aliquet. Aliquam erat volutpat. Vivamus
tristique, nisl eu malesuada semper, libero tortor convallis elit, a
scelerisque orci nisi lacinia turpis. In lacinia ultrices
volutpat. Proin ultrices luctus tellus, in placerat eros tincidunt
id. Ut varius iaculis quam in consequat. Nulla nec orci est, sit amet
pellentesque nisl. Mauris non cursus lectus. Praesent placerat leo vel
erat gravida lacinia. Donec vehicula consectetur lectus vitae
luctus. Praesent nisl justo, laoreet elementum facilisis vel,
tristique ac enim. Etiam vel quam ut quam eleifend
tincidunt. Suspendisse sit amet eros vel elit ullamcorper
laoreet. Etiam venenatis sodales turpis, nec lacinia ligula hendrerit
nec. Nam eu vulputate purus. Quisque facilisis congue metus, sed
imperdiet lorem rhoncus sit amet.

In this paper we have... 
\begin{itemize}
\item Pellentesque vel elit neque, in interdum lacus. Quisque sodales, nunc et luctus convallis, nisl dui luctus dui, at congue urna velit a; 
\item At sit amet sapien a risus dapibus sagittis. Cras sed ultricies erat. Donec id metus sed urna lacinia convallis vel sed enim. 
\item Nisi libero, ornare vel bibendum eu, sollicitudin sed leo. Cras tincidunt aliquet ultricies. Cras pretium velit leo, in malesuada. 
\item Duis sagittis ultricies interdum. Proin sit amet sem nec metus feugiat pharetra.
\end{itemize}

Aliquam ac metus nec odio tempus pharetra sed nec diam. Sed eget arcu
nulla. Etiam elementum ultrices ligula, at iaculis libero feugiat
bibendum. Suspendisse potenti. Nam pharetra adipiscing
euismod. Quisque imperdiet dignissim odio, sed volutpat justo
tincidunt eu. Nunc vehicula pharetra suscipit. Integer aliquet pretium
ipsum vel ultrices. Nam rutrum nibh ac quam pulvinar molestie.

Sed sed ipsum diam. In risus tortor, sagittis eu auctor in, varius in
dui. Mauris a nunc ut ligula ullamcorper tincidunt. Nunc aliquam eros
ac risus pellentesque aliquam. Phasellus augue velit, varius at
porttitor sit amet, pretium eget felis. Ut mollis tellus elementum
magna porttitor rutrum. Etiam blandit leo eget est consectetur
imperdiet. Quisque et diam nec orci vulputate varius vitae id sapien.

\bibliographystyle{mnras}
\bibliography{tester_mnras}

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

\end{document}