### 2019-11-07 tex/other_tex/Ham17_clqs.tex

\begin{table*}
\begin{tabular}{ l rrrr rrrr rrrr rrrr rrrr rr|}
\hline
%  \multicolumn{1}{|c|}{sdss\_name} &
 % \multicolumn{1}{c|}{f1450} &
 % \multicolumn{1}{c|}{$alpha\_civ$} &
 % \multicolumn{1}{c|}{$alpha\_nv$} &
 % \multicolumn{1}{c|}{$alpha\_all$} &
 % \multicolumn{1}{c|}{$alpha\_allc$} &
 % \multicolumn{1}{c|}{rew} &
 % \multicolumn{1}{c|}{rewc} &
 % \multicolumn{1}{c|}{reww} &
 % \multicolumn{1}{c|}{fwhm} &
 % \multicolumn{1}{c|}{fwhme} &
 % \multicolumn{1}{c|}{fwhmc} &
 % \multicolumn{1}{c|}{fwhmw} &
 % \multicolumn{1}{c|}{$\sigma$} &
 % \multicolumn{1}{c|}{peak} &
 % \multicolumn{1}{c|}{peaksnr} &
 % \multicolumn{1}{c|}{rat} &
 % \multicolumn{1}{c|}{asy} &
 % \multicolumn{1}{c|}{kt75} &
 % \multicolumn{1}{c|}{kt80} &
 % \multicolumn{1}{c|}{rew\_di} &
 % \multicolumn{1}{c|}{rewe\_di} &
 % \multicolumn{1}{c|}{fwhm\_di} &
\hline 
  Plate-Fiber-MJD & $f$1450 & $\alpha_{\civ}$ & $\alpha_{\rm all}$      & $\alpha_{\rm allc}$     & rew                       & rewc   &  reww  & fwhm                 & fwhmc  &  fwhmw  & $\sigma$  & asy       & kt75  & kt80   & rew\_di               & fwhm\_di                     \\
\hline 
  5201-178-55832 & 1.4659    & -3.3547              & -4.95$\pm$0.46 & -4.026$\pm$0.452 & 42.76$\pm$1.96  & 26.72  & 16.03  & 5210$\pm$226 & 4651   & 13444    & 3702          &   0.059  & 0.415 & 0.325 & 41.41$\pm$1.6 & 5372  \\
  6118-720-56189 & 0.5101    & -1.5931              & -1.47$\pm$0.04 & -0.657$\pm$0.004 & 42.18$\pm$5.59  &  6.42   & 35.76  & 2994$\pm$620 & 1458   &   7636    & 3002          & -0.150  & 0.183 & 0.138 & 44.15$\pm$4.2 & 3867  \\
\hline
\end{tabular}
  \caption{
Spectrum is SDSS/BOSS MJD-Plate-FiberID combination. 
\noindent {\tt f1450} = flux in the uncorrected BOSS spectrum at 1450
\AA\ rest (10$^{-17}$ ergs s$^{-1}$ cm$^{-2}$ \AA$^{-1}$) used to
anchor the power law continuum fits beneath \civ\ and \nv , e.g.,
$f_{\lambda} = f_{1450}\, (\lambda /1450{\rm \AA})^{\alpha}$.
\noindent {\tt alpha\_civ} = power law continuum slope
($f_{\lambda}\propto \lambda^{\alpha}$) measured from the BOSS
spectrum on either side of \civ .
\noindent {\tt alpha\_nv} = as above but for \nv .
\noindent {\tt alpha\_all} = power law continuum slope ($f_{\lambda}\propto \lambda^{\alpha}$) between 1350 \AA\ and 2200 \AA\ in the uncorrected BOSS spectrum (with 1$\sigma$ uncertainty). 
\noindent {\tt alpha\_allc} = power law continuum slope ($f_{\lambda}\propto \lambda^{\alpha}$) between 1350 \AA\ and 2200 \AA\ in the flux corrected BOSS spectrum, listed only for quasars with $z_{\rm e} \le 3.4$ (with 1$\sigma$ uncertainty). 
\noindent {\tt rew} = \civ\ REW (\AA ) from the line profile fit (with 1$\sigma$ uncertainty). 
\noindent {\tt rewc} = \civ\ REW (\AA ) for the core Gaussian component.
\noindent {\tt reww} = \civ\ REW (\AA ) for the wing Gaussian component.
\noindent {\tt fhwm} = \civ\ FWHM (km/s) from the line profile fit (with 1$\sigma$ uncertainty). 
\noindent {\tt fwhmc} = \civ\ FWHM (km/s) for the core Gaussian only.
\noindent {\tt fwhmw} = \civ\ FWHM (km/s) for the wing Gaussian only.
\noindent {\tt sigma} = \civ\ velocity dispersion (km/s) measured from the profile fit \citep{Peterson2004}.
\noindent {\tt asy} = \civ\ profile asymmetry index \citep{Marziani1996}.
\noindent {\tt kt75} = \civ\ profile kurtosis 75 index \citep{Marziani1996}.
} 
 \label{tab:Ham17_lines}
\end{table*}